---
title: "Spotting address changes at a glance with jsdiffr"
subtitle: "A tiny HTML diff that makes data cleaning reviewable"
format:
  html:
    toc: false
    embed-resources: true
execute:
  warning: false
  message: false
---

```{r}
#| include: false
library(jsdiffr)
```

```{r}
#| echo: false
#| results: asis
# Inject the package's diff colours + table styling once, then provide a helper
# that pulls just the <table> out of diff_html()'s full HTML document (so it
# embeds cleanly without its <body> styles hijacking the page).
cat(
  "<style>\n", diff_css_default(), "\n",
  ".jsdiff-table{border-collapse:collapse;margin:1em 0;font-size:14px}\n",
  ".jsdiff-table th,.jsdiff-table td{padding:4px 10px;border:1px solid #d0d7de;",
  "vertical-align:top;white-space:pre-wrap;font-family:ui-monospace,monospace}\n",
  ".jsdiff-table th{background:#f6f8fa;text-align:left}\n",
  ".jsdiff-table td:nth-child(1){background:#fff8f8}\n",
  ".jsdiff-table td:nth-child(2){background:#f8fff8}\n",
  ".jsdiff-table td:nth-child(3){background:#fafafa}\n",
  "</style>\n", sep = ""
)

diff_table <- function(...) {
  html <- diff_html(..., view = FALSE)
  regmatches(html, regexpr("<table.*</table>", html))
}
```

If you've ever standardised a column of addresses — expanding `St` to `Street`,
fixing a postcode, normalising `Ave` — you know the hard part isn't the
transformation. It's *reviewing* it. A `before`/`after` table of 5,000 rows is
unreadable; your eye can't find the one character that changed.

`jsdiffr` ships two HTML helpers, **`diff_to_html()`** and **`diff_html()`**,
that turn a diff into colour-coded markup: <span class="jsdiff-removed">red
strike-through</span> for what was removed, <span class="jsdiff-added">green</span>
for what was added. It's a one-liner, and it works inline in Quarto, R Markdown,
or a Shiny app.

## One pair, one line

```{r}
#| results: asis
old <- "12 Baker St, London, NW1 6XE"
new <- "12 Baker Street, London, NW1 6EX"

cat(diff_to_html(diff_words(old, new)))
```

At a glance: `St` became `Street`, and the postcode swapped its last two
characters — a typo you'd *never* catch scanning two text columns.

## A whole column, side by side

`diff_html()` takes two vectors and renders an **Original / Updated / Changes**
table. Pass `view = FALSE` to get the HTML back instead of opening a browser.

```{r}
#| results: asis
original <- c(
  "350 Fifth Avenue, New York, NY 10118",
  "1600 Pennsylvania Ave NW, Washington DC",
  "Flat 4, 22 Acacia Avenue, Manchester M14 5TP"
)
cleaned <- c(
  "350 5th Avenue, New York, NY 10118",
  "1600 Pennsylvania Avenue NW, Washington, DC 20500",
  "Flat 4B, 22 Acacia Ave, Manchester M14 5TP"
)

cat(diff_table(original, cleaned, method = diff_words))
```

Every row tells you exactly what your cleaning step did — and, just as
importantly, what it *didn't* touch.

## Hide the noise you don't care about

Reviewing a `St` → `Street` expansion across thousands of rows? That's expected;
you don't want it lighting up red and green. `ignore_words` renders changes made
up entirely of those words as plain context, so only the *surprising* edits
stand out:

```{r}
#| results: asis
cat(diff_table(
  original, cleaned,
  method       = diff_words,
  ignore_words = c("ave", "avenue", "5th", "fifth")
))
```

Now the abbreviation expansions fade into the background and the genuine
changes — the added ZIP code, `4` → `4B` — are what catch your eye.

## Why it's worth reaching for

- **Reviewable cleaning.** Diffs make an address-standardisation pass auditable
  instead of a black box.
- **Zero setup.** No JS toolchain — `diff_to_html()` returns a string you can
  drop straight into a report or `shiny::renderUI()`.
- **Tunable signal.** `ignore_words` and the `diff_*` granularity (`diff_chars`,
  `diff_words`, `diff_sentences`) let you dial in exactly which changes matter.

Two functions, a couple of vectors, and a column of addresses becomes something
you can actually *read*.
