---
title: "Diffing text and sequences with jsdiffr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Diffing text and sequences with jsdiffr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup}
library(jsdiffr)
```

`jsdiffr` is a faithful R port of the
[jsdiff](https://github.com/kpdecker/jsdiff) JavaScript library. It computes the
difference between two pieces of text (or two sequences) using Myers' O(ND)
algorithm, implemented in C++ for speed on large inputs.

## The basics

Every diff function takes two inputs and returns a `data.table` of *change
objects* with four columns: `value`, `added`, `removed`, and `count`.

```{r}
d <- diff_chars("the quick brown fox", "the quiet brown ox")
d
as.data.frame(d)
```

You can diff at several granularities:

```{r}
diff_words("The quick brown fox", "The slow brown fox")
diff_lines("alpha\nbeta\ngamma\n", "alpha\nBETA\ngamma\n")
diff_sentences("Hello world. Foo bar.", "Hello world. Baz bar.")
```

`diff_arrays()` compares arbitrary sequences; its `value` column is a list of
the sub-sequences for each change:

```{r}
da <- diff_arrays(c("a", "b", "c"), c("a", "c", "d"))
da$value
```

## Printing and colours

Printing a diff in an interactive console shows additions in green and deletions
in red. You can force this on or off:

```{r}
cat(format(diff_words("the cat sat", "the dog sat"), color = FALSE), "\n")
```

## Patches

The full unified-diff patch suite is available:

```{r}
old <- "line one\nline two\nline three\n"
new <- "line one\nline 2\nline three\nline four\n"
patch <- create_patch("example.txt", old, new)
cat(patch)
```

Patches can be parsed, applied, reversed, and re-formatted:

```{r}
identical(apply_patch(old, patch), new)

reversed <- format_patch(reverse_patch(parse_patch(patch)))
identical(apply_patch(new, reversed), old)
```

`apply_patch()` supports fuzzy matching via `fuzz_factor` and automatically
reconciles Unix/Windows line endings.

## HTML output for Shiny / R Markdown

`diff_to_html()` renders a diff as HTML spans you can drop into a Shiny app or
report. Pair it with `diff_css_default()` for styling.

```{r}
html <- diff_to_html(diff_words("the cat sat", "the dog sat"))
substr(html, 1, 120)
```

In a Shiny app:

```r
library(shiny)
ui <- fluidPage(
  tags$head(tags$style(HTML(diff_css_default()))),
  uiOutput("diff")
)
server <- function(input, output) {
  output$diff <- renderUI({
    HTML(diff_to_html(diff_lines(old_text(), new_text())))
  })
}
```

## Converters

Changes can be serialised to XML or to Google's diff-match-patch format:

```{r}
convert_changes_to_xml(diff_chars("abc", "abd"))
str(convert_changes_to_dmp(diff_chars("abc", "abd")))
```

## Differences from jsdiff

* The API uses `snake_case`; `camelCase` aliases (`diffChars`, `createPatch`,
  ...) are provided for users porting JavaScript code.
* There is no asynchronous `callback` mode; the universal `max_edit_length` and
  `timeout` options are supported.
* `diff_words()`'s `intlSegmenter` option is not supported (R has no
  `Intl.Segmenter`).
```
